\chapter{Building and Flashing}
\label{ch:build}

The firmware is built with ESP-IDF; the hardware-free domain core also
builds and is unit-tested on the host machine.

\section{Toolchain}

\begin{table}[htbp]
  \centering
  \begin{tabular}{@{}ll@{}}
    \drhead Item & Choice \\
    \midrule
    Framework        & ESP-IDF v5.5.x (native) \\
    Language          & C++23 (\texttt{-std=gnu++23}) \\
    Error model       & \texttt{std::expected}; exceptions off \\
    Security          & NVS + flash encryption (dev mode); see \texttt{docs/security-flash-nvs.md} \\
    Documentation     & Doxygen (build must pass) \\
    \bottomrule
  \end{tabular}
  \caption{Firmware toolchain.}
  \label{tab:build-toolchain}
\end{table}

\section{Si4684 firmware blobs (local only)}

Proprietary Si4684 images are \textbf{not} in git. Before the first device
build:

\begin{drcode}[Populate blobs (from Software/)]
python3 tools/fetch_si4684_firmware.py --dab-only
python3 tools/fetch_si4684_firmware.py --si46xx-dir /path/to/si46xx_firmware
python3 tools/check_si4684_blobs.py
\end{drcode}

See \texttt{Firmware/Si4684-Firmware/README.md} for procurement options.

\section{Device build}

First flash after enabling encryption (fw~0.8.3+) requires a one-time erase:

\begin{drcode}[Build, flash, monitor]
idf.py set-target esp32s3
idf.py build
idf.py erase-flash flash monitor
\end{drcode}

Production flash-encryption release mode uses \texttt{sdkconfig.defaults.production}
as an overlay --- irreversible on the chip; see the security doc before use.

\section{Host unit tests}

The pure core is tested on the host, with no board attached. On macOS the
tests need a C++23 standard library --- use a Homebrew LLVM (>= 18) or
GCC~14 rather than the system Apple Clang.

\begin{drcode}[Host tests]
cmake -S components/core/test -B build-host \
      -DCMAKE_CXX_COMPILER="$(brew --prefix llvm)/bin/clang++"
cmake --build build-host
ctest --test-dir build-host --output-on-failure
\end{drcode}

Thirteen test executables cover JSON parsing, audio design, station list,
integration service, and broadcast metadata accumulators.

\section{Documentation}

Documentation has enforced checks, run from the \texttt{Software/}
directory:

\begin{enumerate}
  \item \textbf{Doxygen} --- C++ API reference from source doc blocks.
        An undocumented public class, method, or parameter fails the build.
  \item \textbf{Manual sync} --- every public class under an
        \texttt{include/} tree must have a matching
        \texttt{\textbackslash label\{cls:ClassName\}} section in
        \texttt{docs/manual/ch-classes.tex}. Design-level HTTP API
        documentation lives in Chapter~\ref{ch:api}.
  \item \textbf{Si4684 blob policy} --- \texttt{tools/check\_si4684\_blobs.py}
        ensures no proprietary \texttt{.bin} is tracked in git.
\end{enumerate}

\begin{drcode}[Docs (from Software/)]
doxygen Doxyfile
python3 tools/check-manual-sync.py
python3 tools/check_si4684_blobs.py
python3 tools/gzip-www.sh    # after editing components/net/www/index.html
\end{drcode}

\section{Continuous integration}

Every push and pull request to \texttt{main} runs
\texttt{.github/workflows/ci.yml} at the repository root (four parallel
jobs, all from the \texttt{Software/} directory):

\begin{enumerate}
  \item \textbf{Host tests} --- \texttt{cmake} + \texttt{ctest} on
        \texttt{components/core/test} (C++23, \texttt{g++-14} on Ubuntu).
  \item \textbf{Doxygen} --- must exit 0 with an empty
        \texttt{docs/api/doxygen-warnings.log}.
  \item \textbf{Manual sync} --- \texttt{tools/check-manual-sync.py}.
  \item \textbf{Si4684 blobs} --- \texttt{tools/check\_si4684\_blobs.py}.
\end{enumerate}

To rebuild the PDF manual (requires a LaTeX installation):

\begin{drcode}[Manual PDF]
cd docs/manual
latexmk -lualatex manual.tex
\end{drcode}

\begin{drcaution}[Keep it green]
All checks are part of the definition of done. A firmware change that
adds or modifies a public class, a REST endpoint, or its behaviour must
update the Doxygen doc blocks, \texttt{ch-classes.tex} (for classes), and
\texttt{ch-api.tex} (for HTTP) in the same change.
\end{drcaution}

\section{Hardware validation}

Automated CI does not attach to a board. When the PCB is available, run the
checklist in \texttt{docs/security-flash-nvs.md} and the HIL items in
\texttt{docs/TODO.md} (section P4).
