% ============================================================
%  DigiRadio — Manual chapter: FSC-BT1035 Bluetooth
% ============================================================

\chapter{FSC-BT1035 Bluetooth Transmitter}
\label{ch:bt1035}

The Feasycom FSC-BT1035 (Qualcomm QCC3056) is the wireless output stage of
DigiRadio: it receives PCM from the ADAU1701 over I\textsuperscript{2}S and
streams Bluetooth audio with aptX, aptX~HD, and aptX~Adaptive. This chapter
documents how the ESP32-S3 controls the module over UART (AT commands with
RTS/CTS), why Line-In mode is mandatory, and how \texttt{bt1035::Bt1035Driver}
implements the bring-up sequence.

\begin{drref}[Hardware context]
Board wiring (UART pins, I\textsuperscript{2}S to the module, flow control)
is in Chapter~\ref{ch:hardware}, Section~\ref{sec:hw-bt1035}. The ADAU1701
master clock and limiter settings that feed the module are in
Chapter~\ref{ch:adau1701} and Chapter~\ref{ch:sigmastudio}.
\end{drref}

\section{Role in the audio chain}
\label{sec:bt1035-role}

The BT1035 is an I\textsuperscript{2}S \textbf{slave} at 48\,kHz: LRCLK and
BCLK come from the ADAU1701; PCM arrives on \texttt{SDATA\_OUT0}
(ADAU MP6 $\rightarrow$ module pin~5). The ESP32 does not process audio
samples for Bluetooth; it only configures the module so the wired path is
accepted and encoded for transmission.

Without firmware init the module may stay in a default mode that ignores the
Line-In from the DSP. The mandatory \texttt{AT+AUXCFG=1} command selects
auxiliary/Line-In input --- omitting it silently breaks the entire wireless
output (Section~\ref{sec:bt1035-linein}).

\section{Control interface}
\label{sec:bt1035-uart}

\subsection{UART parameters}

\begin{table}[htbp]
  \centering
  \begin{tabular}{@{}ll@{}}
    \drhead Setting & Value \\
    \midrule
    Port            & UART2 (not the console UART) \\
    Baud rate       & 115200 \\
    Data format     & 8N1 \\
    Flow control    & Hardware RTS/CTS (required) \\
    Reset           & GPIO active-low pulse at boot \\
    SYS\_CTL        & Held active to enable the module \\
    \bottomrule
  \end{tabular}
  \caption{FSC-BT1035 UART and control lines (\texttt{board\_pins.hpp}).}
  \label{tab:bt1035-uart}
\end{table}

\subsection{Response handling}

Every command expects a module reply containing \texttt{OK} or
\texttt{ERROR}. The driver:

\begin{itemize}
  \item builds lines in the pure core via \texttt{core::buildBt1035AtLine()};
  \item classifies replies with \texttt{core::parseBt1035AtResponse()};
  \item treats timeouts and unexpected payloads as errors (never ignored).
\end{itemize}

Host tests in \texttt{components/core/test/bt1035\_at\_test.cpp} lock the
init sequence (including \texttt{AT+AUXCFG=1}) and the parser.

\section{Mandatory Line-In mode}
\label{sec:bt1035-linein}

\begin{drcaution}[AT+AUXCFG=1 is not optional]
The documented init sequence must include \texttt{AT+AUXCFG=1} after a
successful \texttt{AT} ping. This tells the QCC3056 firmware to take audio
from the wired I\textsuperscript{2}S/Line-In port (the ADAU1701 output)
rather than an internal source. AGENTS.md and the hardware manual both treat
skipping this step as a production bug.
\end{drcaution}

\section{Boot sequence}
\label{sec:bt1035-boot}

At power-up \texttt{HardwareBootstrap::boot()} runs the Si4684 and ADAU1701
first, applies the saved audio profile, then initialises the BT1035 so the
Line-In path is ready before Wi-Fi starts.

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[
      drstep/.style={draw, rounded corners, minimum width=9.5cm,
        minimum height=0.9cm, align=center, font=\small},
      >={Latex}, node distance=3mm]
    \node[drstep, fill=black!6] (sys) {SYS\_CTL high, RESET\# pulse};
    \node[drstep, fill=black!6, below=of sys] (uart)
      {Install UART2 @ 115200, RTS/CTS};
    \node[drstep, fill=black!8, below=of uart] (at)
      {Send \texttt{AT} --- expect OK};
    \node[drstep, fill=black!10, below=of at] (aux)
      {Send \texttt{AT+AUXCFG=1} --- expect OK (Line-In)};
    \node[drstep, fill=black!6, below=of aux] (done)
      {\texttt{Bt1035Driver::isBooted()} = true};
    \foreach \a/\b in {sys/uart, uart/at, at/aux, aux/done} {
      \draw[->] (\a) -- (\b);
    }
  \end{tikzpicture}
  \caption{FSC-BT1035 AT init on DigiRadio.}
  \label{fig:bt1035-boot}
\end{figure}

Pairing, codec selection, and volume over Bluetooth are handled by the
module's own firmware and NVS; DigiRadio firmware currently implements
only the Line-In bring-up required for the wired audio path.

\section{Software architecture}
\label{sec:bt1035-stack}

\begin{table}[htbp]
  \centering
  \small
  \begin{tabular}{@{}llp{5.8cm}@{}}
    \drhead Layer & Type & Responsibility \\
    \midrule
    Bootstrap     & \texttt{HardwareBootstrap} &
      Constructs driver, calls \texttt{boot()} after ADAU1701 \\
    Driver        & \texttt{bt1035::Bt1035Driver} &
      UART, reset, AT init sequence \\
    Pure core     & \texttt{core::Bt1035AtCommand}, parser &
      Host-testable command strings and OK/ERROR classification \\
    \bottomrule
  \end{tabular}
  \caption{BT1035 control stack (Slice~6--7). Pairing and A2DP status are
    exposed on \texttt{/api/bluetooth/*} via \texttt{BluetoothService}.}
  \label{tab:bt1035-stack}
\end{table}

\section{Supported AT command subset}
\label{sec:bt1035-at}

The firmware enumerates every command it sends. Extending the subset requires
updating \texttt{core::Bt1035AtCommand}, the manual, and a host test.

\begin{table}[htbp]
  \centering
  \begin{tabular}{@{}lll@{}}
    \drhead Enum & Line sent & Purpose \\
    \midrule
    \texttt{Ping}      & \texttt{AT}           & Verify UART link \\
    \texttt{AuxLineIn}   & \texttt{AT+AUXCFG=1}  & Enable Line-In from ADAU \\
    \texttt{PairDiscoverable} & \texttt{AT+PAIR=1} & Enter discoverable mode \\
    \texttt{PairHidden}       & \texttt{AT+PAIR=0} & Leave discoverable mode \\
    \texttt{A2dpStat}         & \texttt{AT+A2DPSTAT} & Read link state \\
    \texttt{A2dpDisconnect}   & \texttt{AT+A2DPDISC} & Release A2DP session \\
    \texttt{QueryName}        & \texttt{AT+NAME}     & Read module friendly name \\
    \texttt{QueryAutoConn}    & \texttt{AT+AUTOCONN} & Read auto-reconnect count \\
    \texttt{QueryPairedList}  & \texttt{AT+PLIST}    & List paired remotes \\
    \bottomrule
  \end{tabular}
  \caption{Enumerated AT commands (\texttt{core::Bt1035AtCommand}). Boot
    uses Ping + AuxLineIn only; pairing commands are runtime.}
  \label{tab:bt1035-at}
\end{table}

\section{Bt1035Driver API}
\label{sec:bt1035-driver}

\texttt{Bt1035Driver} owns UART and GPIO reset. All methods return
\texttt{std::expected<T, Bt1035Error>}.

\begin{table}[htbp]
  \centering
  \begin{tabular}{@{}ll@{}}
    \toprule
    \textbf{Method} & \textbf{Purpose} \\
    \midrule
    \texttt{boot()} & Reset, UART init, run \texttt{bootInitSequence()} \\
    \texttt{isBooted()} & \texttt{true} after Line-In init succeeded \\
    \texttt{sendCommand(cmd)} & Send one typed command, expect OK \\
    \texttt{enterPairingMode()} & \texttt{AT+PAIR=1} \\
    \texttt{leavePairingMode()} & \texttt{AT+PAIR=0} \\
    \texttt{queryA2dpState()} & \texttt{AT+A2DPSTAT}, parse \texttt{+A2DPSTAT=} \\
    \texttt{disconnectA2dp()} & \texttt{AT+A2DPDISC} \\
    \texttt{queryDeviceName()} & \texttt{AT+NAME}, parse \texttt{+NAME=} \\
    \texttt{queryAutoReconnect()} & \texttt{AT+AUTOCONN}, parse \texttt{+AUTOCONN=} \\
    \texttt{setAutoReconnect(times)} & \texttt{AT+AUTOCONN=n} (0--15) \\
    \texttt{queryPairedList()} & \texttt{AT+PLIST}, parse \texttt{+PLIST=} lines \\
    \bottomrule
  \end{tabular}
  \caption{Public driver API.}
  \label{tab:bt1035-api}
\end{table}

\subsection{Error codes}

\begin{table}[htbp]
  \centering
  \small
  \begin{tabular}{@{}ll@{}}
    \drhead \texttt{Bt1035Error} & Typical cause \\
    \midrule
    \texttt{ResetFailed}         & GPIO configuration failure \\
    \texttt{UartInitFailed}      & \texttt{uart\_driver\_install} / pins \\
    \texttt{NotBooted}           & \texttt{sendCommand} before \texttt{boot()} \\
    \texttt{AtTimeout}           & No OK/ERROR within 2\,s \\
    \texttt{AtError}             & Module returned ERROR \\
    \texttt{UnexpectedResponse}  & Unrecognised payload (future use) \\
    \bottomrule
  \end{tabular}
  \caption{Driver error enumeration.}
  \label{tab:bt1035-errors}
\end{table}

\section{Integration at power-up}
\label{sec:bt1035-integration}

\texttt{main/hardware\_bootstrap.cpp} constructs a static
\texttt{Bt1035Driver} and calls \texttt{boot()} after
\texttt{AudioService::loadAndApply()}. Failure returns
\texttt{HardwareBootError::Bt1035BootFailed} and \texttt{app\_main} halts
before network bring-up (fail-closed, same as Si4684/ADAU1701).

Boot order:
\begin{enumerate}
  \item Si4684 \texttt{boot(Dab)} --- tuner image in RAM.
  \item ADAU1701 \texttt{boot()} --- SigmaStudio program in RAM.
  \item \texttt{AudioService::loadAndApply()} --- user mixer/EQ profile.
  \item BT1035 \texttt{boot()} --- Line-In enabled for wireless output.
\end{enumerate}

\section{Typical usage (firmware developer)}
\label{sec:bt1035-usage}

After a successful \texttt{HardwareBootstrap::boot()}, the module is ready;
no further calls are required for basic listening. To re-send Line-In config
after a module reset:

\begin{verbatim}
bt1035::Bt1035Driver& bt = ...;
if (auto r = bt.sendCommand(core::Bt1035AtCommand::AuxLineIn); !r) {
  // handle Bt1035Error
}
\end{verbatim}

\section{Further reading}
\label{sec:bt1035-reading}

\begin{itemize}
  \item Feasycom FSC-BT1035 AT command manual (vendor) --- full command set;
        firmware wraps name, paired list, and auto-reconnect for the Web UI.
  \item Chapter~\ref{ch:hardware} --- pin map and I\textsuperscript{2}S routing.
  \item Chapter~\ref{ch:adau1701} --- DSP output that feeds the module.
  \item \texttt{components/core/test/bt1035\_at\_test.cpp} --- init sequence test.
\end{itemize}
