\chapter{Class Reference}
\label{ch:classes}

This chapter documents the firmware's public classes at the design level:
what each class is responsible for, which collaborators it depends on, its
key invariants, and how it fits the system. It complements --- and does
not duplicate --- the generated API documentation, which carries the exact
method signatures. The HTTP JSON endpoints are documented separately in
Chapter~\ref{ch:api}.

\begin{drnote}[How this chapter grows]
Per the development rules, every public, architecturally-significant class
(the drivers, the application services, and the public domain-core types)
gets a section here, tagged \texttt{\textbackslash label\{cls:ClassName\}},
added in the same change that introduces the class. A tooling check keeps
this chapter in step with the code, so it is always current.
\end{drnote}

The class reference tracks firmware~0.8.5 on \texttt{main}. Public classes
are grouped by layer: domain core, application services, and hardware drivers.

% ------------------------------------------------------------------
% Domain core (Slice 1)
% ------------------------------------------------------------------

\section{FirmwareVersion}\label{cls:FirmwareVersion}
Strong type wrapping the firmware release identifier (e.g.\ \texttt{0.8.3}).
Used by \texttt{HealthStatus} and the \texttt{/api/health} endpoint so
version strings are never passed as bare \texttt{char*} across module
boundaries. Invariant: non-empty at construction.

\section{FrequencyKHz}\label{cls:FrequencyKHz}
Validated FM centre frequency in kilohertz (64\,000–108\,000). Parsed once
from JSON at the HTTP boundary via \texttt{FrequencyKHz::tryFromKhz()};
trusted downstream by \texttt{ITuner}, \texttt{Si4684Driver}, and status DTOs.

\section{HealthStatus}\label{cls:HealthStatus}
Immutable health-check DTO returned by \texttt{GET /api/health}. Built via
\texttt{HealthStatus::ok(FirmwareVersion, CompanionChipStatus, serialNumber)}
in the pure core; JSON serialisation lives in \texttt{serializeHealthStatusJson()}.
The \texttt{serialNumber} field carries the EEPROM-derived serial or
\texttt{unknown} when the EUI-48 read fails. No ESP-IDF headers.

\section{Eui48}\label{cls:Eui48}
Strong type for the factory-programmed six-byte EUI-48 read from the
24AA025E48 at word address \texttt{0xFA} (Microchip DS20001191). Exposes
\texttt{serialNumber()} (12 uppercase hex digits) and \texttt{shortSuffix()}
(last three bytes) for SSID and hostname derivation.

\section{DeviceIdentity}\label{cls:DeviceIdentity}
Per-board identity derived from \texttt{Eui48}: SoftAP SSID
\texttt{DigiRadio-<suffix>}, Bluetooth name, STA hostname
\texttt{digiradio-<suffix>}, and canonical serial. \texttt{unknown()} yields
documented fallbacks (\texttt{DigiRadio-setup}, serial \texttt{unknown}).

\section{IDeviceIdentitySource}\label{cls:IDeviceIdentitySource}
Port for reading \texttt{DeviceIdentity} from board storage. Implemented in
the shell by \texttt{eeprom24aa::Eeprom24aa} on the shared I2C bus.

% ------------------------------------------------------------------
% Network shell (Slice 1)
% ------------------------------------------------------------------

\section{SoftApConfig}\label{cls:SoftApConfig}
Immutable value type holding SoftAP parameters (SSID, channel, station
limit). \texttt{forSsid()} builds setup parameters from
\texttt{DeviceIdentity::softApSsid()}; \texttt{setupDefault()} keeps the
\texttt{DigiRadio-setup} fallback when the EEPROM read fails.

\section{SoftApHost}\label{cls:SoftApHost}
RAII wrapper around the ESP-IDF Wi-Fi stack that starts and stops the
configured SoftAP. Imperative shell; no business logic.

\section{SetupWebServer}\label{cls:SetupWebServer}
Minimal HTTP server: gzipped setup UI and typed JSON REST handlers:
\begin{itemize}
  \item \texttt{GET /api/health}, \texttt{POST /api/wifi}
  \item Tuner: \texttt{/api/tuner/*}
  \item Audio: \texttt{/api/audio/*}
  \item Bluetooth: \texttt{/api/bluetooth/*}
  \item Presets: \texttt{/api/stations/*}
\end{itemize}
JSON parsing and serialisation delegate to the pure core; credentials persist via \texttt{ISecureStore};
tuner via \texttt{tuner::TunerService}; audio via
\texttt{audio::AudioService}; Bluetooth via \texttt{bluetooth::BluetoothService};
presets via \texttt{station::StationService}.

\section{NetBootstrap}\label{cls:NetBootstrap}
Owns network resources for setup or STA mode. Initialises encrypted NVS via
\texttt{secure\_store::initEncryptedStorage()}, then Wi-Fi and the HTTP server.
Wires tuner, audio, Bluetooth, station, and integration services into REST
handlers. Must outlive \texttt{app\_main} for the process lifetime.

% ------------------------------------------------------------------
% Domain core + secure store (Slice 2)
% ------------------------------------------------------------------

\section{Secret}\label{cls:Secret}
Opaque wrapper for sensitive strings. Buffer is zeroised on destruction;
no \texttt{operator<<} and no plaintext serialisation. Shell code uses
\texttt{usePlaintext()} transiently for driver APIs only.

\section{WifiSsid}\label{cls:WifiSsid}
Validated Wi-Fi SSID (1--32 bytes). Parsed once from untrusted JSON at
the HTTP boundary before persistence or driver calls.

\section{WifiCredentials}\label{cls:WifiCredentials}
Domain value pairing \texttt{WifiSsid} with a \texttt{Secret} password.
Open networks use an empty password; WPA-PSK requires 8--63 characters.

\section{ISecureStore}\label{cls:ISecureStore}
Abstract persistence boundary for credentials and preset data at rest.
Wi-Fi credentials, station list JSON, and last-preset index use
\texttt{NvsSecureStore}; audio profiles use \texttt{NvsAudioProfileStore}.
Host tests use fakes; the shell uses NVS backends after
\texttt{initEncryptedStorage()}.

\section{StaClient}\label{cls:StaClient}
RAII STA join helper with an explicit connect timeout. Assumes
\texttt{esp\_wifi\_init()} was already called by \texttt{NetBootstrap}.

\section{NvsSecureStore}\label{cls:NvsSecureStore}
\texttt{ISecureStore} implementation backed by an NVS namespace. Passwords
are stored as NVS strings and never logged. NVS encryption and flash
encryption are enabled in \texttt{sdkconfig.defaults}; initialisation runs in
\texttt{secure\_store::initEncryptedStorage()} before network bring-up.

% ------------------------------------------------------------------
% Hardware drivers (Slice 3)
% ------------------------------------------------------------------

\section{IFirmwareBlobReader}\label{cls:IFirmwareBlobReader}
Abstract streaming reader for firmware blobs embedded in flash. The Si4684
driver uses it to HOST\_LOAD patch and application images in bounded SPI
chunks without allocating the full image on the heap. Host tests use the
same interface over in-memory buffers.

\section{EmbeddedBlobReader}\label{cls:EmbeddedBlobReader}
Non-owning \texttt{IFirmwareBlobReader} over a contiguous byte range (linker
symbols from \texttt{EMBED\_FILES} or test data). Implements offset-based
\texttt{read()} with truncation at end-of-blob.

\section{Si4684EmbeddedImages}\label{cls:Si4684EmbeddedImages}
Owns \texttt{EmbeddedBlobReader} views over \texttt{rom\_patch\_016.bin},
\texttt{dab\_firmware.bin}, and \texttt{fm\_firmware.bin}. Constructed once;
passed to \texttt{Si4684Driver}. \texttt{applicationImage(Si4684Band)} selects
DAB or FM.

\section{Si4684Driver}\label{cls:Si4684Driver}
Full SPI driver for the Si4684 DAB+/FM tuner (Chapter~\ref{ch:si4684}).

\texttt{boot(Si4684Band)} runs AN649 HOST\_LOAD, configures I\textsuperscript{2}S
and band-specific properties, and records \texttt{loadedBand()}. After boot the
driver exposes FM tuning/seek/RSQ/RDS, DAB ensemble tuning, digital service list
fetch, and \texttt{startDabService}. Property access (\texttt{setProperty},
\texttt{setVolume}) and diagnostics (\texttt{getPartInfo}, \texttt{getSysState})
are included. Wrong-band calls return \texttt{Si4684Error::WrongBand}.
Register-level opcodes remain private; see \texttt{Si4684Types.hpp} for status
DTOs. Full DAB/FM tuning guide: Chapter~\ref{ch:si4684}.

\section{Bt1035PairedDevice}\label{cls:Bt1035PairedDevice}
Plain DTO for one \texttt{+PLIST=} line from the FSC-BT1035: paired-record
index (1--8), 12-digit MAC, and optional friendly name. Parsed by
\texttt{core::parseBt1035PairedListResponse()} and serialised on
\texttt{GET /api/bluetooth/paired}.

\section{Bt1035Driver}\label{cls:Bt1035Driver}
UART driver for the FSC-BT1035 (Chapter~\ref{ch:bt1035}). \texttt{boot()}
pulses RESET\#, opens UART2 with RTS/CTS, and runs
\texttt{core::bootInitSequence()} (Ping + \texttt{AT+AUXCFG=3} +
\texttt{AT+I2SCFG=67}). Returns
\texttt{Bt1035Error} on timeout, ERROR response, or UART failure.

\section{Adau1701Driver}\label{cls:Adau1701Driver}
RAII I\textsuperscript{2}C driver for the ADAU1701 SigmaDSP
(Chapter~\ref{ch:adau1701}). \texttt{boot()} asserts RESET\#,
initialises the shared I2C bus, and replays the SigmaStudio export from
\texttt{Firmware/ADAU1701-Firmware/} on every power-up (no EEPROM self-boot
on DigiRadio). Exposes \texttt{i2cBusHandle()} after boot for the shared
24AA025E48 EEPROM. Runtime mixer, EQ, and master volume updates use the
ADAU1701 safeload mechanism via \texttt{applyProfile()} and related methods
(Section~\ref{sec:adau1701-driver}).

\section{Eeprom24aa}\label{cls:Eeprom24aa}
I\textsuperscript{2}C shell driver for the 24AA025E48 identity EEPROM
(Chapter~\ref{ch:hardware}, address \texttt{0x52}). Implements
\texttt{IDeviceIdentitySource}: reads six bytes at word address \texttt{0xFA}
per Microchip DS20001191 and builds \texttt{DeviceIdentity}. Called from
\texttt{HardwareBootstrap} after \texttt{Adau1701Driver::boot()}.

\section{Adau1701Dsp}\label{cls:Adau1701Dsp}
\texttt{core::IDsp} adapter over \texttt{Adau1701Driver}. Maps domain-level
audio control to safeload I\textsuperscript{2}C transactions without exposing
parameter RAM addresses to services (Chapter~\ref{ch:adau1701},
Section~\ref{sec:adau1701-stack}).

\section{RegisterWrite}\label{cls:RegisterWrite}
One SigmaStudio \texttt{SIGMA\_WRITE\_REGISTER\_BLOCK} step: 16-bit address
plus payload bytes. Aggregated by \texttt{DspProgram}.

\section{DspProgram}\label{cls:DspProgram}
Ordered ADAU1701 RAM download script (pure domain). Parsed from a framed
\texttt{DRAD} flash blob or built by \texttt{EmbeddedDspProgramSource}.

\section{IDspProgramSource}\label{cls:IDspProgramSource}
Port supplying \texttt{DspProgram} to \texttt{Adau1701Driver::boot()}.
Implementations: flash partition, embedded export, and fallback composite.

\section{EmbeddedDspProgramSource}\label{cls:EmbeddedDspProgramSource}
Builds the factory \texttt{DspProgram} from \texttt{DigiRadio\_IC\_1.h}
(five writes matching \texttt{default\_download\_IC\_1()}).

\section{FlashDspProgramSource}\label{cls:FlashDspProgramSource}
Reads and parses the \texttt{dsp} flash partition; \texttt{storeBlob()} erases
and writes a validated blob from \texttt{POST /api/dsp/program}.

\section{FallbackDspProgramSource}\label{cls:FallbackDspProgramSource}
Tries \texttt{FlashDspProgramSource} first, then
\texttt{EmbeddedDspProgramSource} when the partition is empty or invalid.

% ------------------------------------------------------------------
% Domain core — audio (Slice 5)
% ------------------------------------------------------------------

\section{GainDb}\label{cls:GainDb}
Validated decibel gain/attenuation ($-96$\,dB to $+12$\,dB). Used for
input volumes, master level, and EQ band gain. Converted to ADAU 8.23
fixpoint via \texttt{gainDbToLinearFixpoint()} before safeload.

\section{FrequencyHz}\label{cls:FrequencyHz}
Validated PEQ centre frequency (20--20\,000\,Hz) for the 48\,kHz
SigmaStudio project. Parsed at the HTTP boundary before coefficient design.

\section{EqBandIndex}\label{cls:EqBandIndex}
Strong index into the six-band \texttt{Param EQ1} module (0--5). Maps to
parameter RAM via \texttt{adau1701::paramAddrEqBandBase()}.

\section{EqProfile}\label{cls:EqProfile}
Six-band parametric EQ snapshot aligned with the SigmaStudio map
(Chapter~\ref{ch:sigmastudio}): band index~0 is the fixed 20\,Hz
high-pass; bands 1--5 are peaking at 100\,Hz--8\,kHz. Designed in the
pure core; peaking bands are safeloaded by \texttt{Adau1701Driver::applyEq()}
(band~0 is skipped).

\section{IDsp}\label{cls:IDsp}
Abstract ADAU1701 control boundary. Defines \texttt{applyProfile()},
\texttt{setInputVolume()}, \texttt{setEqBand()}, and related intent-level
operations without ESP-IDF types.

\section{IAudioProfileStore}\label{cls:IAudioProfileStore}
Persistence boundary for \texttt{AudioProfile} (mixer, EQ, master). Device
implementation: \texttt{NvsAudioProfileStore}; host tests use fakes.

% ------------------------------------------------------------------
% Application services (Slice 4–5)
% ------------------------------------------------------------------

\section{ITuner}\label{cls:ITuner}
Abstract tuner boundary in the pure core. Defines boot, band selection,
status readout, FM/DAB tuning, service list, playback, and volume without
ESP-IDF types. \texttt{si4684::Si4684Tuner} implements it on device; host
tests can use fakes.

\section{TunerService}\label{cls:TunerService}
Application service exposing intent-level tuner operations to HTTP and
future UI. Holds a reference to \texttt{core::ITuner}, tracks last tune
target and volume, and maps driver failures to \texttt{core::TunerError}.

\section{Si4684Tuner}\label{cls:Si4684Tuner}
\texttt{ITuner} adapter over \texttt{Si4684Driver}. Translates domain calls
into SPI commands and maps \texttt{Si4684Error} to \texttt{TunerError}.
Constructed once in \texttt{HardwareBootstrap} alongside the driver.

\section{AudioService}\label{cls:AudioService}
Application service for ADAU1701 mixer, EQ, and master volume
(Chapter~\ref{ch:adau1701}, Section~\ref{sec:adau1701-service}). Holds a
reference to \texttt{core::IDsp}, tracks the in-memory \texttt{AudioProfile},
loads from \texttt{IAudioProfileStore} after boot, and applies changes via
safeload. Stereo depth and bass enhance levels are merged into the effective
EQ via \texttt{core::applyEnhancementsToEq()}. Exposed on
\texttt{/api/audio/*} and the web UI Audio section.

\section{EnhanceLevel}\label{cls:EnhanceLevel}
Strong type for 0--100 enhancement intensity (stereo depth and bass boost).
Validated at the HTTP boundary by \texttt{core::parseEnhanceLevelJson()}.

\section{NvsAudioProfileStore}\label{cls:NvsAudioProfileStore}
\texttt{IAudioProfileStore} implementation storing serialised
\texttt{AudioProfile} JSON in NVS namespace \texttt{digiradio}.

\section{StationName}\label{cls:StationName}
Strong type for a preset label (1--32 bytes). Parsed at the HTTP boundary
via \texttt{StationName::tryFrom()}.

\section{BroadcastLabel}\label{cls:BroadcastLabel}
Trimmed on-air label from RDS or DAB chip byte buffers (up to 128 bytes).
Built by \texttt{BroadcastLabel::tryFromChipBytes()} in the pure core.

\section{RdsMetadataAccumulator}\label{cls:RdsMetadataAccumulator}
Stateful RDS decoder assembling program service name and radiotext from
raw block~A--D snapshots returned by the Si4684 driver.

\section{DabDynamicLabelAccumulator}\label{cls:DabDynamicLabelAccumulator}
Reassembles segmented DAB DLS payloads into one dynamic label for
\texttt{TunerStatus}.

\section{PresetSlot}\label{cls:PresetSlot}
Optional hardware preset button index (1--20). Validated by
\texttt{PresetSlot::tryFrom()}.

\section{Station}\label{cls:Station}
Immutable preset value: name, band (DAB/FM), tune coordinates, optional
preset slot. FM presets carry \texttt{FrequencyKHz}; DAB presets carry
ensemble index and optional service/component ids for playback.

\section{StationList}\label{cls:StationList}
Pure-domain ordered collection (max 20 entries) with add/remove/move and
duplicate tune-target rejection. Persisted as JSON through
\texttt{ISecureStore}.

\section{StationService}\label{cls:StationService}
Application service loading/saving presets from NVS and recalling them via
\texttt{TunerService}. Exposed on \texttt{/api/stations/*} and the Presets
web UI section.

\section{IntegrationService}\label{cls:IntegrationService}
End-to-end orchestration at boot and on preset recall: loads the station
list, reapplies the saved \texttt{AudioProfile} after tune, and persists
the last preset index (NVS key \texttt{last\_preset}). Used by
\texttt{app\_main} and \texttt{POST /api/stations/tune}.

\section{BluetoothService}\label{cls:BluetoothService}
Application service for BT1035 pairing and A2DP status
(Chapter~\ref{ch:bt1035}). Delegates to \texttt{Bt1035Driver}; tracks
whether discoverable mode was requested. Exposed on
\texttt{/api/bluetooth/*}.

\section{OtaService}\label{cls:OtaService}
Application service wrapping \texttt{esp\_ota\_ops}: streams firmware into
the inactive OTA slot, validates the app descriptor via
\texttt{core::validateOtaAppDescriptor()}, and exposes
\texttt{confirmBoot()} for rollback cancellation after a healthy network boot.

% ------------------------------------------------------------------
