\chapter{Introduction}
\label{ch:introduction}

DigiRadio is an open-source, high-fidelity digital radio receiver. It
receives DAB+ and FM broadcasts, processes the audio through a dedicated
signal processor, and streams the result over Bluetooth using a
high-resolution codec. Firmware~0.8.3 on \texttt{main} provides encrypted
storage, a tabbed configuration web UI, and the full REST API documented
in Chapter~\ref{ch:api}. The whole project --- hardware and firmware --- is
released as open source for the maker and audio community to study,
build, and improve.

\section{What DigiRadio is}

Most DAB+/FM hobby projects stop at basic reception. DigiRadio adds two
things that lift it into hi-fi territory: a real audio DSP stage for
equalisation and mixing, and a premium-codec Bluetooth output. The result
is a genuine wireless audio source rather than a bench demo, and, being
fully documented, a practical reference for anyone learning multi-chip
audio design or RF-aware PCB layout.

\begin{drkey}[At a glance]
A four-chip design --- Si4684 tuner, ADAU1701 DSP, ESP32-S3 host, and an
FSC-BT1035 Bluetooth transmitter --- on a six-layer board, powered from
USB-C, configured through an elegant web interface.
\end{drkey}

\section{Who this manual is for}

This manual documents the system for someone building, flashing, or
extending DigiRadio: the hardware at a block level
(Chapter~\ref{ch:hardware}), the firmware architecture
(Chapter~\ref{ch:firmware}), dedicated companion-chip guides for the
Si4684 tuner (Chapter~\ref{ch:si4684}) and ADAU1701 DSP
(Chapter~\ref{ch:adau1701}), FSC-BT1035 Bluetooth
(Chapter~\ref{ch:bt1035}), the HTTP JSON API exposed by the web UI
(Chapter~\ref{ch:api}), the per-class design reference that grows with
the code (Chapter~\ref{ch:classes}), and how to build and flash
(Chapter~\ref{ch:build}). Exact C++ signatures are generated by Doxygen
from the source; this manual explains design, behaviour, and the
reasoning behind it.

\section{Licensing summary}

The hardware is licensed under CERN-OHL-S v2 and the firmware under
Apache-2.0; full details are in Chapter~\ref{ch:licensing}. The project
repository is at \url{https://github.com/manvalan/DigiRadio}.
