\chapter{Licensing}
\label{ch:licensing}

DigiRadio is dual-licensed, with hardware and firmware under separate but
compatible open-source licences.

\section{Hardware}

The hardware --- schematics, PCB layout, Gerbers, and bill of materials
--- is licensed under the \textbf{CERN Open Hardware Licence Version~2,
Strongly Reciprocal (CERN-OHL-S v2)}. Anyone who modifies and distributes
the hardware must release their changes under the same terms.

\section{Firmware}

The firmware is licensed under the \textbf{Apache License 2.0}. It grants
patent rights explicitly and pairs naturally with the ESP-IDF ecosystem,
which is Apache-licensed as well. Every source file carries the Apache
header with an SPDX identifier.

\section{Documentation and project pages}

Where a hosting platform does not offer the CERN-OHL-S licence in its
selector, the closest share-alike option is used for that page, with the
authoritative licence stated as CERN-OHL-S v2. The definitive licensing
is always the one declared in the project repository.

\section{Third-party components}

The design integrates third-party silicon (Silicon Labs, Analog Devices,
Espressif, Feasycom/Qualcomm) whose datasheets, firmware images, and DSP
tools remain subject to their respective vendors' terms. Those terms are
unaffected by the DigiRadio licences and must be observed when
redistributing vendor-supplied binaries.

\begin{drnote}[Manufacturing partner]
Prototype fabrication and assembly are supported by PCBWay. Their
contribution is acknowledged in the project documentation and on the
board.
\end{drnote}
